\chapter{Constructing a Lattice}
\label{c:construct-lat}

%---------------------------------------------------------------------------------------------------
\section{Switches}
\label{s:switch}

A \vn{switch} is like an enumerated value except that 


A \vn{switch} is a switch category identifier name, conventionally with the word \vn{Switch} at the end, 
with a set of possible values. Switches are like enums without the associated integer. 
The advantage of switches is that a given switch value can be used for different switch groups. 

For example, 

How to create switches...





%---------------------------------------------------------------------------------------------------
\section{Defining a Lattice Element}
\label{s:ele.def}

The \julia language itself is used as the basis for constructing lattices. Other simulation programs
have similarly utilized the underlying programming language for constructing lattices\cite{merlin++,xsuite},
but this is in marked contrast to such programs as MAD\cite{mad}, Elegant\cite{elegant}, and the 
original \bmad\cite{bmad-orig}. 

Chapter~\sref{c:ele} gives a list of elements defined by \Bmad. Elements are defined using the \vn{@ele}
macro. The general syntax is:
\begin{example}
  @ele eleName = eleType(param1 = val1, param2 = val2, ...)
\end{example}
where \vn{eleName} is the name of the element, \vn{eleType} is the type of element, \vn{param1}, \vn{param2},
etc. are parameter names and \vn{val1}, \vn{val2}, etc. are the parameter values.
Example:
\begin{example}
  @ele qf = Quadrupole(len = 0.6, K1 = 0.370)
\end{example}
The \vn{@ele} macro will construct a \julia variable with the name \vn{eleName}. Additionally the element
that this variable references will also hold \vn{eleName} as the name of the element.

To copy an element use the \vn{deepcopy} constructor.

%---------------------------------------------------------------------------------------------------
\section{Defining a Lattice Element Type}
\label{s:ele.type}

All lattice element types like \vn{Quadrupole}, \vn{Marker}, etc. are subtypes of the abstract type
\vn{Ele}. To construct a new type, use the \vn{\@construct_ele_type} macro. Example:
\begin{example}
  @construct_ele_type MyEleType
\end{example}

%---------------------------------------------------------------------------------------------------
\section{Lattice Element Internals}
\label{s:ele.inside}

All element types have a single component called \vn{pdict} (``parameter dict'') which is of
type \vn{Dict\{Symbol,Any\}}. Using a \vn{Dict} has advantages and disadvantages. The advantage is
that an element is not restricted as to what can be stored in it. The disadvantage is that it is not
type stable (\sref{s:type.stable}). This is generally acceptable when lattices are constructed but
is undesirable during tracking. To regain type stability during tracking, element parameters are
put into immutable structs called \vn{element parameter} groups 
and these structs are stored in \vn{pdict}. During tracking, the tracking
code can access element parameters via the struct which makes the code type stable as will be illustrated below.

The \vn{element parameter} group structures are all subtypes of the abstract type \vn{EleParameterGroup}.
For example, the \vn{LengthGroup} holds the length and s-positions of the element:
\begin{example}
  @kwdef struct LengthGroup <: EleParameterGroup
    L::Float64 = 0
    s::Float64 = 0
    s_downstream::Float64 = 0
  end
\end{example}
The \vn{\@kwdef} macro automatically defines a keyword-based constructor for \vn{LengthGroup}. 
When a parameter group is stored in an element's \vn{pdict}, the key will be the symbol associated
with the struct which in this case is \vn{:LengthGroup}. For example, an element's length can be
accessed via \vn{ele.pdict[:LengthGroup].L}. 


\etcetc...



